BarnebyLives: an R package to create herbarium specimen labels and clean spreadsheets

### 2025-10-25 LabelStyles/skeleton-default.Rmd
label changes - check diff on vignettes

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(dpi = 300) 
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r load data, eval = T}
library(BarnebyLives)
local <- file.path(path.expand('~'), 'Documents', 'assoRted', 'Barneby_Lives_dev', 'LabelStyles')
record <- read.csv(file.path(local, 'SoS-ExampleCollection.csv')) |>
  dplyr::filter(Collection_number == params$Collection_number)

@@

\begin{center}
\textbf{\Large `r writer(record[['Project_name']])`}
\end{center}
\uppercase{`r record[['Family']]`}

\begingroup
\normalsize
`r writer(paste(record[['Genus']], record[['Epithet']]), italics = TRUE)` `r writer(record[['Binomial_authority']])` `r writer(record[['Infrarank']])` `r writer(record[['Infraspecies']], italics = TRUE)` `r writer(record[['Infraspecific_authority']])`  
\endgroup

\vspace{\fill}
\begingroup
\small
`r record[['Gen']]`. `r writer(record[['Site']])` `r record[['latitude_dd']]` `r record[['longitude_dd']]` (`r record[['Datum']]` `r writer(record[['Coordinate_uncertainty']])`).

`r writer(record[['Habitat']])`. `r record[['physical_environ']]`  
Veg.: `r species_font(record[['Vegetation']])` 
`r associates_writer(record[['Associates']], full = TRUE)`
\endgroup

\vspace{\fill}
\begingroup
`r collection_writer(record)`
`r writer_fide(record)`
\endgroup

\begingroup
\scriptsize
`r writer(record[['Research_notes']])`  
\endgroup

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(dpi = 300) 
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r load data, eval = T, echo= F}
library(BarnebyLives)

local <- file.path('~', 'Documents', 'assoRted', 'Barneby_Lives_dev', 'LabelStyles')

record <- read.csv(file.path(local, 'SoS-ExampleCollection.csv')) |>
  sf::st_drop_geometry() |>
  data.frame() |>
  dplyr::filter(Collection_number == params$Collection_number)

@@

\begin{center}
\textbf{\Large `r writer(record[['Project_name']])`}
\end{center}
\begin{flushright}
\uppercase{`r record[['Family']]`}
\end{flushright}
\begingroup
\large

`r writer(paste(record[['Genus']], record[['Epithet']]), italics = TRUE)` `r writer(record[['Binomial_authority']])` `r writer(record[['Infrarank']])` `r writer(record[['Infraspecies']], italics = TRUE)` `r writer(record[['Infraspecific_authority']])`\
\endgroup

\vspace{\fill}
\begingroup
\normalsize

`r record[['Gen']]`. `r writer(record[['Site']])` `r record[['latitude_dd']]` `r record[['longitude_dd']]` (`r record[['Datum']]` `r writer(record[['Coordinate_uncertainty']])`).

`r writer(record[['Habitat']])`. `r record[['physical_environ']]`

Vegetation: `r species_font(record[['Vegetation']])`

`r associates_writer(record[['Associates']], full = TRUE)` \endgroup

\vspace{\fill}
\begingroup

`r collection_writer(record, date=FALSE)`

\begin{flushright}
`r record[['Date_digital_text']]`
\end{flushright}
\endgroup
\begingroup
\small
`r writer(record[['Notes']])`
\endgroup

### 2025-10-25 manuscript/BarnebyLives An R package to create and qc herbarium labels.rmd
label changes - check diff on vignettes

--- 
title: 'BarnebyLives: an R package to create herbarium specimen labels and clean spreadsheets'
author:  |
    | Reed Clark Benkendorf$^1$$^,$$^2$^[Author for Correspondence: rbenkendorf@chicagobotanic.org], Jeremie B. Fant$^1$$^,$$^2$
    |  $^1$Chicago Botanic Garden, 1000 Lake Cook Road, Glencoe, Illinois 60022, USA  
    |  $^2$Plant Biology and Conservation, Northwestern University, Evanston, Illinois 60208, USA  
abstract: |
  \noindent
  **Premise:** Accessioning herbarium specimens is labor-intensive, yet remains vital for research in ecology, evolution, and conservation. As institutional support for herbaria declines, efficient tools are needed to streamline this process. BarnebyLives was developed to assist collectors by supplementing collection notes, verifying taxonomic data, conducting quality checks, generating labels, and submitting digital records.  
  **Methods and Results:** It integrates geospatial data from U.S. government sources to provide jurisdictional and site information, and checks taxonomic names using in-house spell checkers, IPNI author standards, and Kew’s Plants of the World Online. Optional features include generating Google Maps driving directions. The tool outputs data in tabular and spatial formats for review before producing LaTeX-based labels and shipping manifests.     
  **Conclusions:** BarnebyLives improves data accuracy, ensures up-to-date taxonomy, and significantly reduces the time and effort required to accession herbarium specimens.        
keywords: |
  herbarium, natural history museum, collections, geospatial, R, LaTeX   
output: 
  pdf_document: default
  toc: no
  word_document: default
csl: "../citations/american-journal-of-botany.csl"
bibliography: ../citations/citations.bib
link-citations: yes
fig_caption: yes
always_allow_html: yes
header-includes:
  - \usepackage{endfloat}
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}
  - \linenumbers
  - \usepackage{rotating}
---

>>>>

--- 
title: 'BarnebyLives: an R package to create herbarium specimen labels and clean spreadsheets'
author:  |
    |  $^1$Chicago Botanic Garden, 1000 Lake Cook Road, Glencoe, Illinois 60022, USA  
    |  $^2$Plant Biology and Conservation, Northwestern University, Evanston, Illinois 60208, USA  
    | Reed Clark Benkendorf$^1$$^,$$^2$^[Author for Correspondence: rbenkendorf@chicagobotanic.org], Jeremie B. Fant$^1$$^,$$^2$
abstract: |
  \noindent
  **Premise:** Accessioning herbarium specimens is labor-intensive, yet remains vital for research in ecology, evolution, and conservation. As institutional support for herbaria declines, efficient tools are needed to streamline this process. BarnebyLives was developed to assist collectors by supplementing collection notes, verifying taxonomic data, conducting quality checks, generating labels, and submitting digital records.  
  **Methods and Results:** It integrates geospatial data from U.S. government sources to provide jurisdictional and site information, and checks taxonomic names using in-house spell checkers, IPNI author standards, and Kew’s Plants of the World Online. Optional features include generating Google Maps driving directions. The tool outputs data in tabular and spatial formats for review before producing LaTeX-based labels and shipping manifests.     
  **Conclusions:** BarnebyLives improves data accuracy, ensures up-to-date taxonomy, and significantly reduces the time and effort required to accession herbarium specimens.        
keywords: |
  herbarium, natural history museum, collections, geospatial, R, LaTeX   
output: 
  word_document: default
  toc: no
  pdf_document: default
csl: "../citations/american-journal-of-botany.csl"
bibliography: ../citations/citations.bib
link-citations: yes
fig_caption: yes
always_allow_html: yes
header-includes:
  - \usepackage{endfloat}
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}
  - \linenumbers
  - \usepackage{rotating}
---
